---
title: "COVID-19 in the CA Bay Area"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(whisker)

# read in COVID-19 data
bay <- read_csv("covid19_bayarea.csv")
bay$date <- as.Date(bay$date, format = "%Y-%m-%d")

# read in population data
pop <- read_csv("pop_bayarea.csv")

# add county populations to COVID-19 data
# normalize infected total by 100K residents
bay <- inner_join(bay, pop) %>%
  mutate(infected_total_norm = infected_total*100000/population)

# ac_all <- filter(bay, county == "Alameda County" | county == "Berkeley")
ac <- filter(bay, county == "Alameda")
cc <- filter(bay, county == "Contra Costa")
mc <- filter(bay, county == "Marin")
mo <- filter(bay, county == "Monterey")
nc <- filter(bay, county == "Napa")
sf <- filter(bay, county == "San Francisco")
sm <- filter(bay, county == "San Mateo")
scc <- filter(bay, county == "Santa Clara")
scz <- filter(bay, county == "Santa Cruz")
sol <- filter(bay, county == "Solano")
son <- filter(bay, county == "Sonoma")

# Combine data for plots
# Alameda County and City of Berkeley have separate numbers, but combined for now
bay_for_plot <- rbind(ac, cc, mc, mo, nc, sf, sm, scc, scz, sol, son)

# Whisker template for total cases (reusing for diff counties)
template_tot <-
'
{{county_abbv}}_tot <- ggplot({{county_abbv}}, aes(x = date,
                                 y = infected_total,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Total cases: </b>", infected_total, "<br>",
                                               "<b>Additional cases today: </b>", infected_diff),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_point(color = "{{county_color}}") +
  geom_line(color = "{{county_color}}")

ggplotly({{county_abbv}}_tot, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
'

# Whisker template for new cases (reusing for diff counties)
template_new <-
'
{{county_abbv}}_new <- ggplot({{county_abbv}}, aes(x = date,
                                 y = infected_diff,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Additional cases today: </b>", infected_diff),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_col(fill = "{{county_color}}")

ggplotly({{county_abbv}}_new, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
'

# Whisker template for deaths (reusing for diff counties)
template_deaths <-
'
{{county_abbv}}_deaths <- ggplot({{county_abbv}}, aes(x = date,
                                    y = death_total)) +
  xlab("") +
  ylab("Deaths") +
  theme_minimal(base_size = 14) +
  geom_point(data = filter({{county_abbv}}, death_diff > 0),
             aes(text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                               "<b>Total deaths: </b>", death_total, "<br>",
                               "<b>Additional deaths today: </b>", death_diff, "<br>",
                               "<b>Details: </b><br>  ", note),
                 group = 1),
             color = "gray50") +
  geom_line(color = "gray50")

ggplotly({{county_abbv}}_deaths, tooltip = "text") %>%
  config(displayModeBar = FALSE) %>%
  layout(hoverlabel = list(align = "left"))
'

# Data for plots
ac_data <- list(county_abbv = "ac", county_color = "#a6cee3")
cc_data <- list(county_abbv = "cc", county_color = "#1f78b4")
mc_data <- list(county_abbv = "mc", county_color = "#b2df8a")
mo_data <- list(county_abbv = "mo", county_color = "#33a02c")
nc_data <- list(county_abbv = "nc", county_color = "#fb9a99")
sf_data <- list(county_abbv = "sf", county_color = "#e31a1c")
sm_data <- list(county_abbv = "sm", county_color = "#fdbf6f")
scc_data <- list(county_abbv = "scc", county_color = "#ff7f00")
scz_data <- list(county_abbv = "scz", county_color = "#cab2d6")
sol_data <- list(county_abbv = "sol", county_color = "#6a3d9a")
son_data <- list(county_abbv = "son", county_color = "#ffff99")

```

All Counties
=====================================

Normalized data = COVID-19 cases per 100K county residents (based on 2019 population estimates provided by [U.S. Census Bureau QuickFacts](https://www.census.gov/quickfacts/fact/table/US/PST045219))

Column {.tabset}
-------------------------------------
### Original Data, Linear Scale
```{r}
bay_total_chart <- ggplot(bay_for_plot, aes(x = date,
                                   y = infected_total,
                                   color = county,
                                   text = paste0("<b>County: </b>", county, "<br>",
                                                 "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases: </b>", infected_total, "<br>",
                                                 "<b>Additional cases today: </b>", infected_diff),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_total_chart, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Original Data, Log Scale

```{r}
bay_total_log <- ggplot(bay_for_plot, aes(x = date,
                                 y = infected_total,
                                 color = county,
                                 text = paste0("<b>County: </b>", county, "<br>",
                                               "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Total cases: </b>", infected_total, "<br>",
                                               "<b>Additional cases today: </b>", infected_diff),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  scale_y_continuous(trans = "log10")  +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_total_log, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Normalized Data, Linear Scale

```{r}
bay_norm_total_chart <- ggplot(bay_for_plot, aes(x = date,
                                   y = infected_total_norm,
                                   color = county,
                                   text = paste0("<b>County: </b>", county, "<br>",
                                                 "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases (per 100K residents): </b>",
                                                 format(round(infected_total_norm, 2), nsmall = 2)),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases per 100,000 county residents") +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_norm_total_chart, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Normalized Data, Log Scale

```{r}
bay_norm_total_log <- ggplot(bay_for_plot, aes(x = date,
                                   y = infected_total_norm,
                                   color = county,
                                   text = paste0("<b>County: </b>", county, "<br>",
                                                 "<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases (per 100K residents): </b>",
                                                 format(round(infected_total_norm, 2), nsmall = 2)),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases per 100,000 county residents") +
  scale_y_continuous(trans = "log10")  +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Paired",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "Marin",
                            "Monterey",
                            "Napa",
                            "San Francisco",
                            "San Mateo",
                            "Santa Clara",
                            "Santa Cruz",
                            "Solano",
                            "Sonoma"))

ggplotly(bay_norm_total_log, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

Alameda
=====================================

### Confirmed Cases
```{r}
ac_tot_plot <- whisker.render(template_tot,ac_data)
eval(parse(text=ac_tot_plot))
```

### Additional cases by day

```{r}
ac_new_plot <- whisker.render(template_new,ac_data)
eval(parse(text=ac_new_plot))
```

### Deaths
```{r}
ac_deaths_plot <- whisker.render(template_deaths,ac_data)
eval(parse(text=ac_deaths_plot))
```

Contra Costa
=====================================

### Confirmed Cases
```{r}
cc_tot_plot <- whisker.render(template_tot,cc_data)
eval(parse(text=cc_tot_plot))
```

### Additional cases by day

```{r}
cc_new_plot <- whisker.render(template_new,cc_data)
eval(parse(text=cc_new_plot))
```

### Deaths
```{r}
cc_deaths_plot <- whisker.render(template_deaths,cc_data)
eval(parse(text=cc_deaths_plot))
```

Marin
=====================================

### Confirmed Cases
```{r}
mc_tot_plot <- whisker.render(template_tot,mc_data)
eval(parse(text=mc_tot_plot))
```

### Additional cases by day

```{r}
mc_new_plot <- whisker.render(template_new,mc_data)
eval(parse(text=mc_new_plot))
```

### Deaths
```{r}
mc_deaths_plot <- whisker.render(template_deaths,mc_data)
eval(parse(text=mc_deaths_plot))
```

Monterey
=====================================

### Confirmed Cases
```{r}
mo_tot_plot <- whisker.render(template_tot,mo_data)
eval(parse(text=mo_tot_plot))
```

### Additional cases by day

```{r}
mo_new_plot <- whisker.render(template_new,mo_data)
eval(parse(text=mo_new_plot))
```

### Deaths
```{r}
mo_deaths_plot <- whisker.render(template_deaths,mo_data)
eval(parse(text=mo_deaths_plot))
```

Napa
=====================================

### Confirmed Cases
```{r}
nc_tot_plot <- whisker.render(template_tot,nc_data)
eval(parse(text=nc_tot_plot))
```

### Additional cases by day

```{r}
nc_new_plot <- whisker.render(template_new,nc_data)
eval(parse(text=nc_new_plot))
```

### Deaths
```{r}
nc_deaths_plot <- whisker.render(template_deaths,nc_data)
eval(parse(text=nc_deaths_plot))
```

San Francisco
=====================================

### Confirmed Cases
```{r}
sf_tot_plot <- whisker.render(template_tot,sf_data)
eval(parse(text=sf_tot_plot))
```

### Additional cases by day

```{r}
sf_new_plot <- whisker.render(template_new,sf_data)
eval(parse(text=sf_new_plot))
```

### Deaths
```{r}
sf_deaths_plot <- whisker.render(template_deaths,sf_data)
eval(parse(text=sf_deaths_plot))
```

San Mateo
=====================================

### Confirmed Cases
```{r}
sm_tot_plot <- whisker.render(template_tot,sm_data)
eval(parse(text=sm_tot_plot))
```

### Additional cases by day

```{r}
sm_new_plot <- whisker.render(template_new,sm_data)
eval(parse(text=sm_new_plot))
```

### Deaths
```{r}
sm_deaths_plot <- whisker.render(template_deaths,sm_data)
eval(parse(text=sm_deaths_plot))
```

Santa Clara
=====================================

### Confirmed Cases
```{r}
scc_tot_plot <- whisker.render(template_tot,scc_data)
eval(parse(text=scc_tot_plot))
```

### Additional cases by day

```{r}
scc_new_plot <- whisker.render(template_new,scc_data)
eval(parse(text=scc_new_plot))
```

### Deaths
```{r}
scc_deaths_plot <- whisker.render(template_deaths,scc_data)
eval(parse(text=scc_deaths_plot))
```

Santa Cruz
=====================================

### Confirmed Cases
```{r}
scz_tot_plot <- whisker.render(template_tot,scz_data)
eval(parse(text=scz_tot_plot))
```

### Additional cases by day

```{r}
scz_new_plot <- whisker.render(template_new,scz_data)
eval(parse(text=scz_new_plot))
```

### Deaths
```{r}
scz_deaths_plot <- whisker.render(template_deaths,scz_data)
eval(parse(text=scz_deaths_plot))
```

Solano
=====================================

### Confirmed Cases
```{r}
sol_tot_plot <- whisker.render(template_tot,sol_data)
eval(parse(text=sol_tot_plot))
```

### Additional cases by day

```{r}
sol_new_plot <- whisker.render(template_new,sol_data)
eval(parse(text=sol_new_plot))
```

### Deaths
```{r}
sol_deaths_plot <- whisker.render(template_deaths,sol_data)
eval(parse(text=sol_deaths_plot))
```

Sonoma
=====================================

### Confirmed Cases
```{r}
son_tot_plot <- whisker.render(template_tot,son_data)
eval(parse(text=son_tot_plot))
```

### Additional cases by day

```{r}
son_new_plot <- whisker.render(template_new,son_data)
eval(parse(text=son_new_plot))
```

### Deaths
```{r}
son_deaths_plot <- whisker.render(template_deaths,son_data)
eval(parse(text=son_deaths_plot))
```

Sources
=====================================

By [Jack J. Lee](https://www.jackjleescience.com)

Files available on [Github](https://github.com/jackjlee/covid19_bayarea)

**All Counties**

* Cases tracked by [NBC Bay Area](https://www.nbcbayarea.com/news/coronavirus/list-of-coronavirus-cases-in-the-bay-area/2248581/)
* Data from the [LA Times coronavirus project](https://www.latimes.com/projects/california-coronavirus-cases-tracking-outbreak/) saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.latimes.com/projects/california-coronavirus-cases-tracking-outbreak/)
* Data from the [SF Chronicle coronavirus map](https://projects.sfchronicle.com/2020/coronavirus-map/
)

**Alameda**

* Alameda County and the City of Berkeley have separate public health departments
* Alameda County
  - [Dashboard](https://ac-hcsa.maps.arcgis.com/apps/opsdashboard/index.html#/1e0ac4385cbe4cc1bffe2cf7f8e7f0d9)
  - [COVID-19 page](http://www.acphd.org/2019-ncov.aspx) from the Alameda County Public Health Department
  - Past numbers from the Alameda County Public Health Department, saved by the [Wayback Machine](https://web.archive.org/web/*/http://www.acphd.org/2019-ncov.aspx)
  - [Press releases](http://www.acphd.org/2019-ncov/press-releases.aspx) by the Alameda County Public Health Department
  - Coronavirus reporting by the [Mercury News](https://www.mercurynews.com/tag/coronavirus/)
  - Last updated using data from May 17, 2020
* City of Berkeley
  - [Dashboard](https://www.cityofberkeley.info/covid19-data/)
  - [COVID-19 page](https://www.cityofberkeley.info/covid19/) from the Public Health Division
  - [Coronavirus news](https://www.cityofberkeley.info/coronavirus-news/) provided by the Public Health Division
  - Reporting by [Berkeleyside](https://www.berkeleyside.com/)
  - Last updated using data from May 17, 2020

**Contra Costa**

* [Current numbers](https://www.coronavirus.cchealth.org/) from the Contra Costa Health Services
* Past numbers from the Contra Costa Health Services, saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.coronavirus.cchealth.org/)
* Coronavirus reporting by the [Mercury News](https://www.mercurynews.com/tag/coronavirus/)
* Last updated using data from May 17, 2020

**Marin**

* [Coronavirus status updates](https://coronavirus.marinhhs.org/surveillance) provided daily by Marin Health and Human Services
* Last updated using data from May 17, 2020

**Monterey**

* [COVID-19 media page](https://www.co.monterey.ca.us/government/departments-a-h/health/diseases/2019-novel-coronavirus-2019-ncov/covid-19-related-press-releases) maintained by the County of Monterey Health Department ("Daily Situational Report")
* Last updated using data from May 15, 2020 (numbers from May 13)

**Napa**

* [Coronavirus (COVID-19) updates](https://www.countyofnapa.org/2770/Situation-Updates) provided by Napa County
* [Situation updates](https://www.countyofnapa.org/2770/Situation-Updates) provided by Napa County Public Health
* Last updated using data from May 16, 2020

**San Francisco**

* [Current numbers](https://www.sfdph.org/dph/alerts/coronavirus.asp) from the San Francisco Department of Public Health
* Past numbers from the San Francisco Department of Public Health, saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.sfdph.org/dph/alerts/coronavirus.asp)
* [Press releases](https://www.sfdph.org/dph/alerts/coronavirus-pressreleases.asp) by the San Francisco Department of Public Health
* Last updated using data from May 17, 2020

**San Mateo**

* [Dashboard](https://www.smchealth.org/post/san-mateo-county-covid-19-data-1)
* [Coronavirus page](https://www.smchealth.org/coronavirus) by San Mateo County Health
* Past numbers from San Mateo County Health, saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.smchealth.org/coronavirus)
* Last updated using data from May 15, 2020

**Santa Clara**

* [Coronavirus (COVID-19) Data Dashboard](https://www.sccgov.org/sites/covid19/Pages/dashboard.aspx) from the County of Santa Clara Public Health Department
* [Tweets](https://twitter.com/HealthySCC) and [press releases](https://www.sccgov.org/sites/phd/news/Pages/newsroom.aspx) by the County of Santa Clara Public Health Department
* Coronavirus reporting by the [Mercury News](https://www.mercurynews.com/tag/coronavirus/)
* Last updated using data from May 17, 2020

**Santa Cruz**

* [Coronavirus page](http://www.santacruzhealth.org/HSAHome/HSADivisions/PublicHealth/CommunicableDiseaseControl/Coronavirus.aspx) maintained by the County of Santa Cruz Health Services Agency
* Previously, numbers on the Health Services Agency page, saved by the [Wayback Machine](https://web.archive.org/web/*/http://www.santacruzhealth.org/HSAHome/HSADivisions/PublicHealth/CommunicableDiseaseControl/Coronavirus.aspx)
* Coronavirus reporting by the [Santa Cruz Sentinel](https://www.santacruzsentinel.com/tag/coronavirus/)
* Last updated using data from May 17, 2020

**Solano**

* [Solano County COVID-19 Dashboard ](https://doitgis.maps.arcgis.com/apps/MapSeries/index.html?appid=055f81e9fe154da5860257e3f2489d67)
* Last updated using data from May 15, 2020

**Sonoma**

* [Coronavirus Cases Dashboard ](https://socoemergency.org/emergency/novel-coronavirus/resources-information/dashboard/) provided by the County of Sonoma
* Last updated using data from May 17, 2020
