---
title: "COVID-19 in the CA Bay Area"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(whisker)

# read in data
bay <-read_csv("covid19_bayarea.csv")
bay$date <- as.Date(bay$date, format = "%Y-%m-%d")


# ac_all <- filter(bay, county == "Alameda County" | county == "Berkeley")
ac <- filter(bay, county == "Alameda")
cc <- filter(bay, county == "Contra Costa")
sf <- filter(bay, county == "San Francisco") 
sm <- filter(bay, county == "San Mateo")
scc <- filter(bay, county == "Santa Clara")
scz <- filter(bay, county == "Santa Cruz")

bay_for_plot <- rbind(ac, cc, sf, sm, scc, scz)

# Whisker template for total cases (reusing for diff counties)
template_tot <-
'
{{county_abbv}}_tot <- ggplot({{county_abbv}}, aes(x = date,
                                 y = infected_total,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Total cases: </b>", infected_total, "<br>",
                                               "<b>Additional cases today: </b>", infected_diff),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_point(color = "{{county_color}}") +
  geom_line(color = "{{county_color}}")

ggplotly({{county_abbv}}_tot, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
'

# Whisker template for new cases (reusing for diff counties)
template_new <-
'
{{county_abbv}}_new <- ggplot({{county_abbv}}, aes(x = date,
                                 y = infected_diff,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Additional cases today: </b>", infected_diff),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_col(fill = "{{county_color}}")

ggplotly({{county_abbv}}_new, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
'

# Whisker template for deaths (reusing for diff counties)
template_deaths <-
'
{{county_abbv}}_deaths <- ggplot({{county_abbv}}, aes(x = date,
                                    y = death_total)) +
  xlab("") +
  ylab("Deaths") +
  theme_minimal(base_size = 14) +
  geom_point(data = filter({{county_abbv}}, death_diff > 0),
             aes(text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                               "<b>Total deaths: </b>", death_total, "<br>",
                               "<b>Additional deaths today: </b>", death_diff, "<br>",
                               "<b>Details: </b><br>  ", note),
                 group = 1),
             color = "gray50") +
  geom_line(color = "gray50")

ggplotly({{county_abbv}}_deaths, tooltip = "text") %>%
  config(displayModeBar = FALSE) %>%
  layout(hoverlabel = list(align = "left"))
'

# Data for plots
ac_data <- list(county_abbv = "ac", county_color = "#e41a1c")
cc_data <- list(county_abbv = "cc", county_color = "#377eb8")
sf_data <- list(county_abbv = "sf", county_color = "#4daf4a")
sm_data <- list(county_abbv = "sm", county_color = "#984ea3")
scc_data <- list(county_abbv = "scc", county_color = "#ff7f00")
scz_data  <- list(county_abbv = "scz", county_color = "#ffff33")

```

All Counties 
===================================== 

**Note:** Work in progress - more counties to come soon

Column {.tabset}
-------------------------------------
### Confirmed Cases - Linear Scale
```{r}
bay_total_chart <- ggplot(bay_for_plot, aes(x = date,
                                   y = infected_total,
                                   color = county,
                                   text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                                 "<b>Total cases: </b>", infected_total, "<br>",
                                                 "<b>Additional cases today: </b>", infected_diff),
                                   group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Set1",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "San Francisco",
                            "San Mateo", 
                            "Santa Clara", 
                            "Santa Cruz"))

ggplotly(bay_total_chart, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Confirmed Cases - Log Scale

```{r}
bay_total_log <- ggplot(bay_for_plot, aes(x = date,
                                 y = infected_total,
                                 color = county,
                                 text = paste0("<b>Date: </b>", format(date, format = "%b %d"), "<br>",
                                               "<b>Total cases: </b>", infected_total, "<br>",
                                               "<b>Additional cases today: </b>", infected_diff),
                                 group = 1)) +
  xlab("") +
  ylab("Confirmed cases") +
  scale_y_continuous(trans = "log10")  +
  theme_minimal(base_size = 14) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Set1",
                   name="",
                   breaks=c("Alameda",
                            "Contra Costa",
                            "San Francisco",
                            "San Mateo", 
                            "Santa Clara", 
                            "Santa Cruz"))

ggplotly(bay_total_log, tooltip = "text") %>%
  config(displayModeBar = FALSE)  %>%
  layout(hoverlabel = list(align = "left"))
```

### Confirmed Cases - Normalized

Coming soon

Alameda
===================================== 

### Confirmed Cases
```{r}
ac_tot_plot <- whisker.render(template_tot,ac_data)
eval(parse(text=ac_tot_plot))
```

### Additional cases by day

```{r}
ac_new_plot <- whisker.render(template_new,ac_data)
eval(parse(text=ac_new_plot))
```

### Deaths
```{r}
ac_deaths_plot <- whisker.render(template_deaths,ac_data)
eval(parse(text=ac_deaths_plot))
```

Contra Costa
===================================== 

### Confirmed Cases
```{r}
cc_tot_plot <- whisker.render(template_tot,cc_data)
eval(parse(text=cc_tot_plot))
```

### Additional cases by day

```{r}
cc_new_plot <- whisker.render(template_new,cc_data)
eval(parse(text=cc_new_plot))
```

### Deaths
```{r}
cc_deaths_plot <- whisker.render(template_deaths,cc_data)
eval(parse(text=cc_deaths_plot))
```

San Francisco
===================================== 

### Confirmed Cases
```{r}
sf_tot_plot <- whisker.render(template_tot,sf_data)
eval(parse(text=sf_tot_plot))
```

### Additional cases by day

```{r}
sf_new_plot <- whisker.render(template_new,sf_data)
eval(parse(text=sf_new_plot))
```

### Deaths
```{r}
sf_deaths_plot <- whisker.render(template_deaths,sf_data)
eval(parse(text=sf_deaths_plot))
```

San Mateo
===================================== 

### Confirmed Cases
```{r}
sm_tot_plot <- whisker.render(template_tot,sm_data)
eval(parse(text=sm_tot_plot))
```

### Additional cases by day

```{r}
sm_new_plot <- whisker.render(template_new,sm_data)
eval(parse(text=sm_new_plot))
```

### Deaths
```{r}
sm_deaths_plot <- whisker.render(template_deaths,sm_data)
eval(parse(text=sm_deaths_plot))
```

Santa Clara
===================================== 

### Confirmed Cases
```{r}
scc_tot_plot <- whisker.render(template_tot,scc_data)
eval(parse(text=scc_tot_plot))
```

### Additional cases by day

```{r}
scc_new_plot <- whisker.render(template_new,scc_data)
eval(parse(text=scc_new_plot))
```

### Deaths
```{r}
scc_deaths_plot <- whisker.render(template_deaths,scc_data)
eval(parse(text=scc_deaths_plot))
```

Santa Cruz
===================================== 

### Confirmed Cases
```{r}
scz_tot_plot <- whisker.render(template_tot,scz_data)
eval(parse(text=scz_tot_plot))
```

### Additional cases by day

```{r}
scz_new_plot <- whisker.render(template_new,scz_data)
eval(parse(text=scz_new_plot))
```

### Deaths

None reported

Sources
===================================== 

By [Jack J. Lee](https://www.jackjleescience.com)

Files available on [Github](https://github.com/jackjlee/covid19_bayarea)

**All Counties**

* Cases tracked by [NBC Bay Area](https://www.nbcbayarea.com/news/coronavirus/list-of-coronavirus-cases-in-the-bay-area/2248581/)
* Data from the [LA Times coronavirus project](https://www.latimes.com/projects/california-coronavirus-cases-tracking-outbreak/) saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.latimes.com/projects/california-coronavirus-cases-tracking-outbreak/)
* Data from the [SF Chronicle coronavirus map](https://projects.sfchronicle.com/2020/coronavirus-map/
)

**Alameda**

* Alameda County and the City of Berkeley have separate public health departments
* Alameda County
  - [Current numbers](http://www.acphd.org/2019-ncov.aspx) from the Alameda County Public Health Department
  - Past numbers from the Alameda County Public Health Department, saved by the [Wayback Machine](https://web.archive.org/web/*/http://www.acphd.org/2019-ncov.aspx)
  - [Press releases](http://www.acphd.org/2019-ncov/press-releases.aspx) by the Alameda County Public Health Department
  - Coronavirus reporting by the [Mercury News](https://www.mercurynews.com/tag/coronavirus/)
  - Last updated using data from March 26, 2020
* City of Berkeley
  - [Current numbers](https://www.cityofberkeley.info/covid19/) from the Public Health Division
  - [Coronavirus news](https://www.cityofberkeley.info/coronavirus-news/) provided by the Public Health Division
  - Reporting by [Berkeleyside](https://www.berkeleyside.com/)
  - Last updated using data from March 26, 2020

**Contra Costa**

* [Current numbers](https://www.coronavirus.cchealth.org/) from the Contra Costa Health Services
* Past numbers from the Contra Costa Health Services, saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.coronavirus.cchealth.org/)
* Coronavirus reporting by the [Mercury News](https://www.mercurynews.com/tag/coronavirus/)
* Last updated using data from March 26, 2020

**San Francisco**

* [Press releases](https://www.sfdph.org/dph/alerts/coronavirus-pressreleases.asp) by the San Francisco Department of Public Health
* [Current numbers](https://www.sfdph.org/dph/alerts/coronavirus.asp) from the San Francisco Department of Public Health
* Past numbers from the San Francisco Department of Public Health, saved by the [Wayback Machine](https://web.archive.org/web/*/https://www.sfdph.org/dph/alerts/coronavirus.asp)
* Last updated using data from March 26, 2020

**Santa Clara**

* [Tweets](https://twitter.com/HealthySCC) and [press releases](https://www.sccgov.org/sites/phd/news/Pages/newsroom.aspx) by the County of Santa Clara Public Health Department
* [Current numbers](https://www.sccgov.org/sites/phd/DiseaseInformation/novel-coronavirus/Pages/home.aspx) from the  County of Santa Clara Public Health Department
* Coronavirus reporting by the [Mercury News](https://www.mercurynews.com/tag/coronavirus/)
* Last updated using data from March 26, 2020

**Santa Cruz**

* Coronavirus reporting by the [Santa Cruz Sentinel](https://www.santacruzsentinel.com/tag/coronavirus/)
* [Current numbers](http://www.santacruzhealth.org/HSAHome/HSADivisions/PublicHealth.aspx) from the County of Santa Cruz Health Services Agency
* Past numbers from the County of Santa Cruz Health Services Agency, saved by the [Wayback Machine](https://web.archive.org/web/*/http://www.santacruzhealth.org/HSAHome/HSADivisions/PublicHealth/CommunicableDiseaseControl/Coronavirus.aspx)
* Last updated using data from March 26, 2020
